---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-"
)
```

# detrange

## Introduction 

`detrange` estimates detection range (DR) from multiple stations within a passive acoustic telemetry array using range testing data collected in the field. DR is defined by Kessel et al. (2014) as

>"... the relationship between detection probability and the distance between the receiver and tag...presented graphically in the form of a logistic curve of detection probability."

Given a modeled DR, it is possible to estimate the distance at which a target detection efficiency (DE) occurs. DE is defined by Brownscombe et al (2019) as 

>"[t]he number of acoustic transmitter detections effectively logged by an acoustic receiver in a given time period, expressed as a percentage (or proportion) of total potential detections based on transmission rate."

Following recommendations from Brownscombe et al (2019) and Huveneers et al. (2016), it can be useful to estimate the midpoint of the DR (i.e. distance at 50% DE) in order to place sentinel tags at a sample of receivers to measure variation in DE over time.

Under the hood, `detrange` uses a Bayesian mixed-effects logistic linear regression model implemented in the JAGS language. Station is treated as a random effect (slope and intercept) and Distance is the sole covariate.

## Demonstration

### Data

`detrange` expects data typical of detection range testing. Mandatory columns include:  

- `Station` (factor)  
- `Distance` (numeric)  
- `Detects` (integer)  
- `Pings` (integer)

`Pings` is the expected number of detections. An example dataset `range_test` is included for reference.

```{r data}
library(detrange)

### view example dataset
data <- range_test
data
```

### Analysis
To estimate detection range, use `dr_analyse()`. The `priors` argument allows the user to change the default priors (in JAGS language), `de_target` specifies the desired detection efficiency to estimate distance at and `nthin` can be adjusted to improve model convergence. Note a single prior can be changed and the default priors will remain for the rest.
```{r}
### set target DE to 50%
de_target <- 0.5
### replace default prior for bIntercept
prior_bIntercept <- list(bIntercept = "dnorm(10, 5^-2)") 

analysis <- dr_analyse(data, de_target = de_target, priors = prior_bIntercept, nthin = 1L)
```
The output of `dr_analyse()` is an object of class `mbr`. It can be manipulated using any functions within the [mbr package](https://github.com/poissonconsulting/mbr). For convenience, `detrange` provides some functions to summarise/visualise the analysis.

```{r}
### plot predicted values
dr_plot_predicted(analysis, de_target = de_target)
```

```{r}
### get midpoint estimates
dr_analysis_midpoint(analysis)
```

```{r}
### coefficient table
dr_analysis_coef(analysis)
```

## Code of Conduct

Please note that the detrange project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/0/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.

## Installation

Install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("Freshwater-Fish-Ecology-Laboratory/detrange")
```
