---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-"
)
```

# detrange

<!-- badges: start -->
<!-- badges: end -->

## Introduction 

`detrange` estimates detection range of receivers within passive acoustic telemetry arrays. Under the hood is a Bayesian mixed-effects logistic linear regression model, where each station has a random intercept and slope. 

## Demonstration

### Range test data
Range test data must include columns `Station`, `Distance`, `Detects` and `Pings`.

```{r data}
library(detrange)
library(mbr)
library(jmbr)

data <- detrange::range_test
data
dr_plot_observed(data)
```

### Analysis
Model the detection range with `dr_analyse`.
```{r}
analysis <- detrange::dr_analyse(data, nthin = 1L)

### coefficient table
dr_analysis_coef(analysis)

### get midpoint estimates
dr_analysis_midpoint(analysis)
```
### Plot
Plot predicted detection range, midpoints and confidence intervals. 
```{r}
### plot predicted values
dr_plot_predicted(analysis)
```

## Code of Conduct

Please note that the detrange project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/0/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.

## Installation

Install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("Freshwater-Fish-Ecology-Laboratory/detrange")
```
