---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-"
)
```

# detrange 

## Introduction 

#### Detection Range and Detection Efficiency
`detrange` estimates detection range (DR) from multiple stations within a passive acoustic telemetry array using range testing data collected in the field. DR is defined by Kessel et al. (2014) as

>"... the relationship between detection probability and the distance between the receiver and tag...presented graphically in the form of a logistic curve of detection probability."

Given a modeled DR, it is possible to estimate the distance at which a target detection efficiency (DE) occurs. DE is defined by Brownscombe et al (2019) as 

>"[t]he number of acoustic transmitter detections effectively logged by an acoustic receiver in a given time period, expressed as a percentage (or proportion) of total potential detections based on transmission rate."

Following recommendations from Brownscombe et al (2019) and Huveneers et al. (2016), it can be useful to estimate the midpoint of the DR (i.e. distance at 50% DE) in order to place sentinel tags at a sample of receivers to measure variation in DE over time.

#### The modelling approach
Under the hood, `detrange` uses JAGS software and the [rjags](https://cran.r-project.org/web/packages/rjags/rjags.pdf) R package to implement a Bayesian generalized linear model with logit link and binomial response distribution. If there are 5 or more stations, the model is fit as a generalized linear mixed-effects model with random intercept and slope for each Station. Otherwise, Station is treated as a fixed effect. 

A benefit of using a Bayesian approach is that uncertainty can be quantifed for estimates of the distance at which a specified DE occurs. In Bayesian lingo, the uncertainty of this derived parameter can be estimated with the posterior distributions of other parameters in the model. 

Another benefit is the ability to incorporate prior information. By default, the priors used in the model are non-informative. However, the user may set custom priors, e.g., if prior information about realistic detection range in a given system is known or if data is limited. 

## Demonstration

### Data

`detrange` expects data typical of detection range testing. Mandatory columns include:  

- `Station` (factor)  
- `Distance` (numeric)  
- `Detects` (integer)  
- `Pings` (integer)

`Pings` is the expected number of detections and `Detects` is the observed number of detections over the duration of the range testing time period at a given distance. An example dataset `range_test` is included for reference.

```{r data}
library(detrange)

### view example dataset
data <- range_test
head(data)
```

### Analysis
To estimate detection range, use `dr_analyse()`
```{r}
analysis <- dr_analyse(data)
```
Check model convergence with `dr_glance()`. Try adjusting the `nthin` argument in `dr_analyse()` to improve convergence. Higher `nthin` generates more sample iterations.

```{r}
dr_glance(analysis)
```

Look at model coefficient estimates `dr_coef()`

```{r}
coef <- dr_coef(analysis, conf_level = 0.89)
head(coef)
```

Predict distance at which a target level of detection efficiency occurs with `dr_distance_at_de()`
```{r}
midpoint <- dr_distance_at_de(analysis, de_target = 0.5, conf_level = 0.8)
midpoint
```


Predict detection efficiency at a sequence of distances with `dr_predict()`
```{r}
predicted <- dr_predict(analysis, distance_seq = seq(0, 1000, 100)) 
head(predicted)
```

Plot results with `dr_plot_predicted()`
```{r}
### coefficient table
# dr_analysis_coef(analysis)
```

### How to do more
The output of `dr_analyse()` is an object of s3 class `dr_model`. It contains a list with two elements:  
1. `analysis$model` - the model object of class `jags` created by `rjags::jags.model()`    
1. `analysis$samples` - the MCMC samples generated from`rjags::jags.samples()` and converted to `mcmcr` class


These objects are the raw materials for any further exploration or analysis. For example, view trace and density plots with `plot(analysis$samples)`.

See [mcmcr](https://github.com/poissonconsulting/mcmcr) and [mcmcderive](https://github.com/poissonconsulting/mcmcderive) for working with `mcmcr` objects, or convert samples to an object of class `mcmc.list`, e,g, with `coda::as.mcmc.list` for working with the [coda](https://github.com/cran/coda) R package. 

## Code of Conduct

Please note that the detrange project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/0/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.

## Installation

Install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("Freshwater-Fish-Ecology-Laboratory/detrange")
```
